Modular Model-Checking of a Byzantine Fault-Tolerant Protocol

### 2017-03-02 paper/paper.tex
Fix last typo and rewording

\begin{abstract}
With proof techniques like IC3 and $k$-induction, model-checking scales further
than ever before. Still, fault-tolerant distributed systems are particularly
challenging to model-check given their large state spaces and
non-determinism. The typical approach to controlling complexity is to construct
ad-hoc abstractions of faults, message-passing, and behaviors. However, these
abstractions come at the price of divorcing the model from its implementation
and making refactoring difficult. In this work, we present a model for
fault-tolerant distributed system verification that combines ideas from the
literature including calendar automata, symbolic fault injection, and abstract
transition systems, and then use it to model-checking various implementations of
the Hybrid Oral Messages algorithm that differ in the fault model, timing model,
and local node behavior. We show that despite being implementation-level models,
the verifications are scalable and modular, insofar as isolated changes to an
implementation require isolated changes to the model and proofs. This work is
carried out in the SAL model-checker.
\end{abstract}

@@

The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,Lincoln-Rushby,pvs}. If formal verification is to be introduced into the workflow of system designers, though, we need more automated methods that scale for implementation-level models. (Mostly) automated proof techniques are required to reduce the need for specialized verification expertise. We also need programmatic verification of \emph{implementations}. System designers create software and hardware implementations to test, simulate, and deploy. Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former~\cite{paxos}, particularly if those abstractions are ad-hoc and system specific. Furthermore, as implementations are modified to explore the design space, it is easy for the formal model and the implementation to become inconsistent, so the verification is no longer about the system deployed.

>>>>

\begin{abstract}
With proof techniques like IC3 and $k$-induction, model-checking scales further
than ever before. Still, fault-tolerant distributed systems are particularly
challenging to model-check given their large state spaces and
non-determinism. The typical approach to controlling complexity is to construct
ad-hoc abstractions of faults, message-passing, and behaviors. However, these
abstractions come at the price of divorcing the model from its implementation
and making refactoring difficult. In this work, we present a model for
fault-tolerant distributed system verification that combines ideas from the
literature including calendar automata, symbolic fault injection, and abstract
transition systems, and then use it to model-check various implementations of
the Hybrid Oral Messages algorithm that differ in the fault model, timing model,
and local node behavior. We show that despite being implementation-level models,
the verifications are scalable and modular, insofar as isolated changes to an
implementation require isolated changes to the model and proofs. This work is
carried out in the SAL model-checker.
\end{abstract}

@@

The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation. Historically, it has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,Lincoln-Rushby,pvs}. If formal verification is to be introduced into the workflow of system designers, though, we need more automated methods that scale for implementation-level models. (Mostly) automated proof techniques are required to reduce the need for specialized verification expertise. We also need programmatic verification of \emph{implementations}. System designers create software and hardware implementations to test, simulate, and deploy. Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former~\cite{paxos}, particularly if those abstractions are ad-hoc and system specific. Furthermore, as implementations are modified to explore the design space, it is easy for the formal model and the implementation to become inconsistent, so the verification is no longer about the system deployed.
